### 2021-10-22 text/outline/outline.tex
add bv decomp

\begin{itemize}
\item \begin{proposition}
The Bayes-optimal threshold $c_\textrm{bayes}$ is a critical value of the limiting bias $b$ (holding other parameters fixed; Figure \ref{thresholding_theoretical}a, all panels).
\end{proposition}
For certain values of $\beta^g_0$ and $\beta^g_1$, $c_\textrm{bayes}$ is a maximizer of the bias (Figure \ref{thresholding_theoretical}a, left); for other values of $\beta^g_0$ and $\beta^g_1$, $c_\textrm{bayes}$ is a minimizer of the bias (Figure \ref{thresholding_theoretical}a, right).
\item
\begin{proposition}
The limit of the bias in $c$ is $1/2$, i.e. 
$$\lim_{c \to \infty} b(\beta^g_0, \beta^g_1, c, 1/2) = 1/2.$$
\end{proposition}
 In other words, we always can set the threshold to a large number and achieve a bias of $1/2$ (Figure \ref{thresholding_theoretical}a, all panels). This fact establishes an upper bound on the bias of thresholded regression (under optimal threshold selection).
 \item
 \begin{proposition}
 For $\beta^g_1 < 2 \Phi(3/4) \approx 1.35$, $c = c_\textrm{bayes}$ yields a \textit{larger} bias than $c = \infty$ (Figure \ref{thresholding_theoretical}a, left); for $\beta^g_1 > 2 \Phi(3/4)$, $c_\textrm{bayes}$ yields a \textit{smaller} bias than $c = \infty$ (Figure \ref{thresholding_theoretical}a, right); finally, for $\beta^g_1 = 2 \Phi(3/4)$, $c = c_\textrm{bayes}$ and $c = \infty$ both yield a bias of $1/2$ (Figure \ref{thresholding_theoretical}a, middle).
 \end{proposition}
\item \begin{proposition}
Limiting bias decreases monotonically as $\beta_1^g$ increases (Figure \ref{thresholding_theoretical}a; fix threshold and scan panels from left to right).
\end{proposition}
This is consistent with the intuition that the problem becomes easier as the gRNA mixture distribution becomes increasingly well-separated.
\item 
\begin{proposition} For all $\beta^g_0, \beta^g_1,$ and $c \in \R,$ the bias is strictly greater than $0$ and less than $2$ (Figure \ref{thresholding_theoretical}a, all panels). 
\end{proposition}
Therefore, the thresholding method suffers from attenuation bias.
\end{itemize}

@@

Next, we generalize the expression for bias when the threshold is large to arbitrary $\pi \in [0,1/2]$:
\begin{proposition}
The following relationship holds:
$$ \lim_{ c \to \infty } b(\beta^g_0, \beta^g_1, c, \pi) = \pi.$$
\end{proposition}
In other words, when the perturbation probability is equal to $\pi$ and the threshold large, the bias is $\pi$. We can understand this result intuitively by considering an extreme example. When $\pi$ is very small (e.g., $\pi = 0.01$), almost all cells are unperturbed. Therefore, if we select a large threshold, we correctly classify nearly all unperturbed cells as unperturbed; on the other hand, the \textit{perturbed} cells that we misclassify as \textit{unperturbed} are swamped in number by the truly unperturbed cells, resulting in a small bias.

@@

% Finally, to shed light on limitations of the large threshold selection strategy, we study the variance of the thresholding estimator. Let $v$ denote the limiting variance of scaled estimator $\sqrt{n} \hat{\beta}^m_1$, i.e.,
% $$v(\beta^g_0, \beta^g_1, c, \pi) = \lim_{n \to \infty} \V(\sqrt{n} \hat{\beta}^m_1).$$ The limiting variance of $\sqrt{n} \hat{\beta}^m_1$ grows without bound as the threshold $c$ tends to infinity: 
% $$ \lim_{c \to \infty} v(\beta^g_0, \beta^g_1, c, \pi) = \infty.$$ Thus, while selecting a large threshold is effective from a bias reduction standpoint when $\pi$ is small, it causes the variance to explode.

To more precisely characterize the bias-variance tradeoff of the thresholding method, we consider a slightly simpler model: we set $\beta^m_0$ and $\beta^g_0$ to zero in (\ref{theoretical_model}) and study the no-intercept least squares estimator
$$ \hat{\beta}^m_1 =  \frac{ \sum_{i=1}^n \hat{p}_i m_i }{ \sum_{i=1}^n \hat{p}_i^2}.$$ We uncover an exact bias-variance tradeoff as a function of the selected threshold (Figure \ref{thresholding_theoretical}c).

\begin{proposition}
The limiting distribution is:
$$\sqrt{n}(\hat{\beta}^m - \beta^m) \xrightarrow{d} N(0, v).$$
\end{proposition}

The bias-variance analysis reveals that the best strategy for minimizing mean-squared error is to select a threshold that induces moderate bias. (A downside of this approach, of course, is that constructing valid confidence intervals becomes more challenging.)

@@

We derive the limit in probability of $\hat{\beta}^m_1$ for the Gaussian model with intercepts (\ref{theoretical_model}). Dividing by $n$ in (\ref{thresh_est_intercepts}), we can express $\hat{\beta}^m_1$ as
$$ \hat{\beta}^m_1 = \frac{ \frac{1}{n} \sum_{i=1}^n ( \hat{p}_i - \overline{\hat{p}_i})(m_i - \overline{m})}{ \frac{1}{n} \sum_{i=1}^n (\hat{p}_i - \overline{\hat{p}})}.$$ By weak LLN,
$$ \hat{\beta}^m_1 \xrightarrow{P} \frac{\textrm{Cov}(\hat{p}_i, m_i)}{\V\left(\hat{p}_i\right)}.$$ To compute this quantity, we first compute several simpler quantities:
\begin{itemize}
\item[1.] Expectation of $m_i$: $\E[m_i] = \beta^m_0 + \beta^m_1\pi.$
\item[2.] Expectation of $\hat{p}_i$: Let $f$ denote the $N(0,1)$ density. Then \begin{multline*}
\E[\hat{p}_i] = \P\left[\hat{p}_i = 1\right] = \P\left[\beta^g_0 + \beta^g_1 + \tau_i \geq c \right] = \\ \textrm{(By LOTP) } \P\left[ \beta^g_0 + \tau_i \geq c \right]\P\left[p_i = 0\right] + \P\left[ \beta^g_0 + \beta^g_1 + \tau_i \geq c \right] \P[p_i = 1] \\ = \P\left[ \tau_i \geq c - \beta^g_0\right](1- \pi) + \P\left[ \tau_i \geq c - \beta^g_1 - \beta^g_0 \right](\pi) \\ = \left(\int_{c - \beta^g_0}^\infty f\right) (1 - \pi) + \left( \int_{c - \beta^g_1 - \beta^g_0}^\infty f \right)(\pi) = \zeta(1-\pi) + \omega \pi.
\end{multline*}
\item[3.] Expectation of $\hat{p}_i p_i$: 
$$\E\left[ \hat{p}_i p_i \right] = \E\left[ \hat{p}_i | p_i = 1 \right] \P\left[ p_i =1 \right] = \P\left[ \beta^g_0 + \beta^g_1 + \tau_i \geq c \right] \pi = \omega \pi.$$
\item[4.] Expectation of $\hat{p}_i m_i$:
\begin{multline*}
\E\left[\hat{p}_i m_i\right] = \E[\hat{p}_i (\beta^m_0 + \beta^m_1 p_i + \ep_i)] = \beta^m_0 \E\left[\hat{p}_i\right] + \beta^m_1 \E\left[\hat{p}_i p_i\right] + \E[\hat{p}_i \ep_i] \\ = \beta^m_0 \E[\hat{p}_i] + \beta^m_1 \omega \pi + \E[\hat{p}_i] \E[\ep_i] = \beta^m_0 \E[\hat{p}_i] + \beta^m_1 \omega \pi.
\end{multline*}
\item[5.] Variance of $\hat{p}_i$: Because $\hat{p}_i$ is binary, we have that $\V[\hat{p}_i] = \E[\hat{p}_i]\left(1 - \E[\hat{p}_i]\right) .$
\item[6.] Covariance of $\hat{p}_i, m_i$:
\begin{multline*}
\textrm{Cov}\left(\hat{p}_i, m_i\right) = \E\left[\hat{p}_i m_i\right] - \E[\hat{p}_i] \E[m_i] = \beta^m_0 \E[\hat{p}_i] + \beta^m_1 \omega \pi - \E[\hat{p}_i]( \beta^m_0 + \beta^m_1 \pi)\\ = \beta^m_1 \omega \pi - \E[\hat{p}_i] \beta_1^m \pi = \beta^m_1 \pi \left( \omega - \E[\hat{p}_i]\right).
\end{multline*}
Combining these expressions, we have that
$$ \hat{\beta}^m_1 \xrightarrow{P} \frac{\beta^m_1 \pi (\omega - \E[\hat{p}_i])}{\E[\hat{p}_i](1 - \E[\hat{p}_i])} = \beta^m_1 \gamma(\beta^g_1, c, \pi).$$
\end{itemize}

@@

>>>>

\begin{itemize}
\item \begin{proposition}
The Bayes-optimal threshold $c_\textrm{bayes}$ is a critical value of the limiting bias $b$ (holding other parameters fixed; Figure \ref{thresholding_theoretical}a, all panels).
\end{proposition}
For certain values of $\beta^g_0$ and $\beta^g_1$, $c_\textrm{bayes}$ is a maximizer of the bias (Figure \ref{thresholding_theoretical}a, left); for other values of $\beta^g_0$ and $\beta^g_1$, $c_\textrm{bayes}$ is a minimizer of the bias (Figure \ref{thresholding_theoretical}a, right).
\item
\begin{proposition}
The limit of the bias in $c$ is $1/2$, i.e. 
$$\lim_{c \to \infty} b(\beta^g_0, \beta^g_1, c, 1/2) = 1/2.$$
\end{proposition}
 In other words, we always can set the threshold to a large number and achieve a bias of $1/2$ (Figure \ref{thresholding_theoretical}a, all panels). This fact establishes an upper bound on the bias of thresholded regression (under optimal threshold selection).
 \item
 \begin{proposition}
 For $\beta^g_1 < 2 \Phi(3/4) \approx 1.35$, $c = c_\textrm{bayes}$ yields a \textit{larger} bias than $c = \infty$ (Figure \ref{thresholding_theoretical}a, left); for $\beta^g_1 > 2 \Phi(3/4)$, $c_\textrm{bayes}$ yields a \textit{smaller} bias than $c = \infty$ (Figure \ref{thresholding_theoretical}a, right); finally, for $\beta^g_1 = 2 \Phi(3/4)$, $c = c_\textrm{bayes}$ and $c = \infty$ both yield a bias of $1/2$ (Figure \ref{thresholding_theoretical}a, middle).
 \end{proposition}
\item \begin{proposition}
Limiting bias decreases monotonically as $\beta_1^g$ increases (Figure \ref{thresholding_theoretical}a; fix threshold and scan panels from left to right).
\end{proposition}
This is consistent with the intuition that the problem becomes easier as the gRNA mixture distribution becomes increasingly well-separated.
\item 
\begin{proposition} For all $\beta^g_0, \beta^g_1,$ and $c \in \R,$ the incurs strict attenuation bias (Figure \ref{thresholding_theoretical}a, all panels).
\end{proposition}
\end{itemize}

@@

Next, we generalize the expression for bias to arbitrary $\pi \in [0,1/2]$:
\begin{proposition}
The following relationship holds:
$$ \lim_{ c \to \infty } b(\beta^g_0, \beta^g_1, c, \pi) = \pi.$$
\end{proposition}
In other words, if the perturbation probability is $\pi$, and if we set the threshold to a large number, then the bias also is $\pi$. We can understand this result intuitively by considering an extreme example. When $\pi$ is very small (e.g., $\pi = 0.01$), almost all cells are unperturbed. Therefore, in selecting a large threshold, we correctly classify nearly all unperturbed cells as unperturbed; on the other hand, the \textit{perturbed} cells that we misclassify as \textit{unperturbed} are swamped in number by the truly unperturbed cells, resulting in a small bias.

@@

To shed light on possible limitations of the large threshold selection strategy, we derive an exact bias-variance decomposition for the thresholding estimator. We consider a slightly simpler, no-intercept version of (\ref{theoretical_model}):
\begin{equation}\label{theoretical_model_no_int}
\begin{cases}
m_i = \beta_m p_i + \ep_i \\
g_i = \beta_g p_i + \tau_i \\
p_i \sim \textrm{Bern}(\pi) \\
\ep_i, \tau_i \sim N(0,1) \\
p_i \indep \tau_i \indep \ep_i,
\end{cases}
\end{equation}
The thresholding estimator $\hat{\beta}_m$ is given by
\begin{equation}\label{thresh_estimator_no_int}
 \hat{\beta}_m = \frac{ \sum_{i=1}^n \hat{p}_i m_i }{\sum_{i=1}^n \hat{p}_i^2}. 
\end{equation}
\begin{proposition}
The limiting distribution of $\hat{\beta}_m$ is
$$\sqrt{n}\left(\hat{\beta}_m - l \right) \xrightarrow{d} N\left(0, v\right).$$
\end{proposition}
The bias decreases and the variance increases as the threshold tends to infinity, consistent with the intuition that a large threshold yields a low misclassification rate at the cost of a small ``effective sample size.'' The best strategy for maximizing estimation accuracy (as quantified by mean squared error) is to select a threshold that induces moderate bias. A downside of this approach, however, is that constructing valid confidence intervals becomes more challenging.

@@

We derive the limit in probability of $\hat{\beta}^m_1$ for the Gaussian model with intercepts (\ref{theoretical_model}). Dividing by $n$ in (\ref{thresh_est_intercepts}), we can express $\hat{\beta}^m_1$ as
$$ \hat{\beta}^m_1 = \frac{ \frac{1}{n} \sum_{i=1}^n ( \hat{p}_i - \overline{\hat{p}_i})(m_i - \overline{m})}{ \frac{1}{n} \sum_{i=1}^n (\hat{p}_i - \overline{\hat{p}})}.$$ By weak LLN,
$$ \hat{\beta}^m_1 \xrightarrow{P} \frac{\textrm{Cov}(\hat{p}_i, m_i)}{\V\left(\hat{p}_i\right)}.$$ To compute this quantity, we first compute several simpler quantities:
\begin{itemize}
\item[1.] Expectation of $m_i$: $\E[m_i] = \beta^m_0 + \beta^m_1\pi.$
\item[2.] Expectation of $\hat{p}_i$: Let $f$ denote the $N(0,1)$ density. Then \begin{multline*}
\E[\hat{p}_i] = \P\left[\hat{p}_i = 1\right] = \P\left[\beta^g_0 + \beta^g_1 p_i + \tau_i \geq c \right] = \\ \textrm{(By LOTP) } \P\left[ \beta^g_0 + \tau_i \geq c \right]\P\left[p_i = 0\right] + \P\left[ \beta^g_0 + \beta^g_1 + \tau_i \geq c \right] \P[p_i = 1] \\ = \P\left[ \tau_i \geq c - \beta^g_0\right](1- \pi) + \P\left[ \tau_i \geq c - \beta^g_1 - \beta^g_0 \right](\pi) \\ = \left(\int_{c - \beta^g_0}^\infty f\right) (1 - \pi) + \left( \int_{c - \beta^g_1 - \beta^g_0}^\infty f \right)(\pi) = \zeta(1-\pi) + \omega \pi.
\end{multline*}
\item[3.] Expectation of $\hat{p}_i p_i$: 
$$\E\left[ \hat{p}_i p_i \right] = \E\left[\hat{p}_i | p_i = 1 \right] \P\left[ p_i =1 \right] = \P\left[ \beta^g_0 + \beta^g_1 + \tau_i \geq c \right] \pi = \omega \pi.$$
\item[4.] Expectation of $\hat{p}_i m_i$:
\begin{multline*}
\E\left[\hat{p}_i m_i\right] = \E[\hat{p}_i (\beta^m_0 + \beta^m_1 p_i + \ep_i)] = \beta^m_0 \E\left[\hat{p}_i\right] + \beta^m_1 \E\left[\hat{p}_i p_i\right] + \E[\hat{p}_i \ep_i] \\ = \beta^m_0 \E[\hat{p}_i] + \beta^m_1 \omega \pi + \E[\hat{p}_i] \E[\ep_i] = \beta^m_0 \E[\hat{p}_i] + \beta^m_1 \omega \pi.
\end{multline*}
\item[5.] Variance of $\hat{p}_i$: Because $\hat{p}_i$ is binary, we have that $\V[\hat{p}_i] = \E[\hat{p}_i]\left(1 - \E[\hat{p}_i]\right) .$
\item[6.] Covariance of $\hat{p}_i, m_i$:
\begin{multline*}
\textrm{Cov}\left(\hat{p}_i, m_i\right) = \E\left[\hat{p}_i m_i\right] - \E[\hat{p}_i] \E[m_i] = \beta^m_0 \E[\hat{p}_i] + \beta^m_1 \omega \pi - \E[\hat{p}_i]( \beta^m_0 + \beta^m_1 \pi)\\ = \beta^m_1 \omega \pi - \E[\hat{p}_i] \beta_1^m \pi = \beta^m_1 \pi \left( \omega - \E[\hat{p}_i]\right).
\end{multline*}
Combining these expressions, we have that
$$ \hat{\beta}^m_1 \xrightarrow{P} \frac{\beta^m_1 \pi (\omega - \E[\hat{p}_i])}{\E[\hat{p}_i](1 - \E[\hat{p}_i])} = \beta^m_1 \gamma(\beta^g_1, c, \pi).$$
\end{itemize}

@@

We prove the bias-variance decomposition for the no-intercept model (\ref{theoretical_model_no_int}). Define $l$ (for ``limit'') by
$$l = \beta_m \left(\frac{\omega \pi}{\zeta(1-\pi) + \omega \pi}\right),$$ where
$$
\begin{cases}
\omega = \int_{c - \beta_g}^\infty f = 1 - \Phi(\beta_g - c) \\
\zeta = \int_{c}^\infty f = 1 - \Phi(-c).
\end{cases}
$$
We have that
\begin{equation*}
\hat{\beta}_m - l = \frac{\sum_{i=1}^n \hat{p}_i m_i}{ \sum_{i=1}^n \hat{p}^2_i} - l = \frac{\sum_{i=1}^n \hat{p}_i m_i}{ \sum_{i=1}^n \hat{p}^2_i} - \frac{l \sum_{i=1}^n \hat{p}_i^2 }{ \sum_{i=1}^n \hat{p}_i^2} \\ = \frac{\sum_{i=1}^n \hat{p}_i(m_i - l \hat{p}_i)}{ \sum_{i=1}^n \hat{p}_i^2}.
\end{equation*}
Therefore,
\begin{equation}\label{bc_decomp_1}
\sqrt{n}(\hat{\beta}_m - l) = \frac{(1/\sqrt{n})\sum_{i=1}^n \hat{p}_i(m_i - l \hat{p}_i)}{(1/n)\sum_{i=1}^n \hat{p}_i^2}.
\end{equation}
Next, we compute the expectation and variance of $\hat{p}_i(m_i - l\hat{p}_i)$. To do so, we first compute several simpler quantities:
\begin{enumerate}
\item Expectation of $\hat{p}_i$: 
\begin{multline*}
\E[\hat{p}_i] = \P(p_i\beta_g + \tau_i \geq c) =  \P(\beta_g + \tau_i \geq c)\pi + \P(\tau_i \geq c)(1-\pi) \\ = \pi \omega + (1-\pi)\zeta.
\end{multline*}
\item Expectation of $\hat{p}_i p_i$: $$\E\left[\hat{p}_i p_i\right] = \E\left[\hat{p}_i | p_i = 1 \right]\P\left[p_i = 1\right] = \omega \pi.$$
\item Expectation of $\hat{p}_i m_i$:
\begin{multline*}
\E[\hat{p}_i m_i] = \E\left[\hat{p}_i(\beta_m p_i + \ep_i)\right] = \E\left[\beta_m \hat{p}_i p_i + \hat{p}_i \ep_i \right] \\ = \beta_m \E\left[ \hat{p}_i p_i \right] + \E[\hat{p}_i]\E[\ep_i] = \beta_m \omega \pi + 0 = \beta_m \omega \pi.
\end{multline*}
\item Expectation of $\hat{p}_i m_i^2$: \begin{multline*}
\E\left[\hat{p}_i m_i^2\right] = \E \left[ \hat{p}_i( \beta_m p_i + \ep_i )^2 \right] = \E\left[ \hat{p}_i \left( \beta_m^2 p_i^2 + 2 \beta_m p_i \ep_i + \ep_i^2 \right)  \right] \\ = \E\left[ \hat{p}_i p_i \beta^2_m + 2 \beta_m p_i \hat{p}_i \ep_i + \hat{p}_i \ep_i^2 \right] = \beta^2_m \E[ \hat{p}_i p_i] + 2 \beta_m \E[p_i\hat{p}_i] \E[\ep_i] + \E[\hat{p}_i] \E[ \ep^2_i ] \\ = \beta^2_m \E[ \hat{p}_i p_i] + \E[\hat{p}_i] = \beta^2_m \omega \pi + \E[ \hat{p}_i]. 
\end{multline*}
\end{enumerate}

Now, we can compute the expectation and variance of $\hat{p}_i(m_i - l\hat{p}_i)$. First,
\begin{multline}\label{bv_decomp_2}
\E\left[\hat{p}_i(m_i - l\hat{p}_i) \right] = \E[\hat{p}_i m_i] - l \E[\hat{p}_i] \\ = \beta_m \omega \pi - \left(\frac{\beta_m \omega \pi}{\zeta (1-\pi) + \omega \pi}\right)[\zeta (1-\pi) + \omega \pi] = 0.
\end{multline}
Additionally,
\begin{multline}\label{bv_decomp_3}
\V\left[\hat{p}_i(m_i - l\hat{p}_i)\right] = \E\left[\hat{p}_i^2(m_i - l\hat{p}_i)^2\right] - (\E\left[ \hat{p}_i(m_i - l\hat{p}_i)\right])^2 \\ = \E\left[ \hat{p}_i m_i^2\right] - 2l \E[m_i\hat{p}_i] +l^2 \E[\hat{p}_i]= \beta^2_m \omega \pi + \E[ \hat{p}_i] -2l \beta_m \omega \pi + l^2 \E[\hat{p}_i] \\ = \beta_m\omega\pi(\beta_m - 2l) + \E[\hat{p}_i](1 + l^2).
\end{multline}
Therefore, by CLT, (\ref{bv_decomp_2}), and (\ref{bv_decomp_3}),
\begin{equation}\label{bv_decomp_4}
(1/\sqrt{n})\sum_{i=1}^n \hat{p}_i(m_i - l \hat{p}_i) \xrightarrow{d} N\left(0, \beta_m\omega\pi(\beta_m - 2l) + \E[\hat{p}_i](1 + l^2) \right).
\end{equation}
Next, by weak LLN,
\begin{equation}\label{bv_decomp_5}
(1/n) \sum_{i=1}^n \hat{p}_i^2 = (1/n) \sum_{i=1}^n \hat{p}_i \xrightarrow{P} \E[\hat{p}_i].
\end{equation}
Finally, by (\ref{bc_decomp_1}), (\ref{bv_decomp_4}), (\ref{bv_decomp_5}), and Slutsky's Theorem,
$$ \sqrt{n}(\hat{\beta}_m - l) \xrightarrow{d} N\left(0, \frac{ \beta_m\omega\pi(\beta_m - 2l) + \E[\hat{p}_i](1 + l^2) }{\left(\E[\hat{p}_i]\right)^2} \right).$$ Thus, for large $n \in \N$, we find that 
$$ \begin{cases}
\E [\hat{\beta}_m] \approx l, \\
\V[\hat{\beta}_m] \approx \left[\beta_m\omega\pi(\beta_m - 2l) + \E[\hat{p}_i](1 + l^2)\right]/[n\E^2[\hat{p}_i]],
\end{cases}$$
completing the bias-variance decomposition.
